---
title: "Project-Patient-Return"
author: "thaque2050"
date: "March 31, 2019"
output: html_document
---

## Define caption font for the tables
```{r results="asis"}
cat("
<style>
caption {
      color: black;
      font-weight: 600;
      font-size: 1.1em;
    }
</style>
")
```
<br>

## Import libraries
```{r,warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(knitr)
library(kableExtra)
library(reshape2)
library(caTools)
library(xgboost)
library(h2o)
library(ranger)
library(class)
library(tree)
library(MASS)
```
<br>

## Exploratory analysis of combined test and train data
<br>

### View training data
```{r,warning=FALSE}
df<-read.csv('Hospitals_Train.csv',nrows = 38221)
print(paste('Number of features are',ncol(df)))
print(paste('Number of observations are',nrow(df)))
str(df)
```
<br>

### View testing data
```{r,warning=FALSE}
df2<-read.csv('Hospitals_Test_X.csv')
print(paste('Number of features are',ncol(df2)))
print(paste('Number of observations are',nrow(df2)))
```
<br>

### Combine testing and training data for exploratory analysis
```{r,warning=FALSE}
df3=data.frame(rbind(df[,1:26],df2))
```
<br>

### Missing values by variable
```{r,warning=FALSE}
df4=df3
df4[df4=='']=NA
df4[df4=='#N/A']=NA
df4[df4=='#VALUE!']=NA

miss_df3_abs=sapply(df4, function(x)sum(is.na(x))/nrow(df4)*100)%>%data.frame()
m_variable=rownames(miss_df3_abs)
miss_df3_abs=data.frame(cbind(m_variable,miss_df3_abs))
colnames(miss_df3_abs)=c('Variable_Name','Missing_Values')

miss_df3_abs%>%ggplot()+
  geom_col(aes(x=reorder(Variable_Name,-Missing_Values),y=Missing_Values),color='blue',fill='blue')+
  xlab('Variable Name')+ylab('Percent of Missing Values')+
  theme(axis.text.x = element_text(angle = 90))+ylim(0,100)+ggtitle('Figure 0: Percent Missing Entries by Feature')
```
<br>


### View consult_in_ed feature
```{r,warning=FALSE}
kable(table(df3$CONSULT_IN_ED,useNA = 'always'),caption = 'Table 1: Frequency of factors in consult_in_ed feature')%>%kable_styling(bootstrap_options = c("striped", "hover"))
```
<br>

### View risk and severity features
```{r,warning=FALSE}
df3%>%group_by(SEVERITY)%>%summarise(count=n())%>%ggplot()+geom_col(aes(x = reorder(SEVERITY,-count),y=count),color='blue',fill='blue')+ylab('Number of Observations')+theme(axis.text.x = element_text(angle = 90))+ggtitle('Figure 1: Frequency of factors in severity feature')

df3%>%group_by(RISK)%>%summarise(count=n())%>%ggplot()+geom_col(aes(x = reorder(RISK,-count),y=count),color='blue',fill='blue')+ylab('Number of Observations')+theme(axis.text.x = element_text(angle = 90))+ggtitle('Figure 2: Frequency of factors in risk feature')

kable(df3%>%group_by(RISK,SEVERITY)%>%summarise(count=n())%>%data.frame(),caption = 'Table 2: Missing values between risk and severity')%>%kable_styling(bootstrap_options = c("striped", "hover"))
```
<br>

### View admit_result feature
```{r,warning=FALSE}
df3%>%group_by(ADMIT_RESULT)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x = reorder(ADMIT_RESULT,-count),y = count),color='blue',fill='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 3: Frequency of factors in admit_result feature')

a=df3%>%filter(ADMIT_RESULT=='')%>%group_by(ADMIT_RESULT,SAME_DAY,ED_RESULT)%>%
  summarise(count=n())%>%data.frame()

kable(a,caption = 'Table 3: Admit_result vs same_day vs ED_RESULT')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
<br>


### View acuity_arr feature
```{r,warning=FALSE}
df3%>%group_by(ACUITY_ARR)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x = reorder(ACUITY_ARR,-count),y = count),color='blue',fill='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 4: Frequency of factors in acuity_arr feature')

a=df3[df3$ACUITY_ARR=='',]%>%mutate(total=n())%>%group_by(RISK,SEVERITY,total)%>%summarise(count=n())%>%
  mutate(proportion=paste(round(count/total*100,digits = 2),'%',sep =''))%>%data.frame()%>%
  dplyr::select(RISK,SEVERITY,count,proportion)

kable(a,caption = 'Table 4: Mapping missing entries in acuity_arr with values in risk/severity')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
<br>

### View same_day feature
```{r,warning=FALSE}
df3%>%group_by(SAME_DAY)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x = reorder(SAME_DAY,-count),y = count),color='blue',fill='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 5: Frequency of factors in same_day feature')
```
<br>

### View consult_order feature
```{r,warning=FALSE}
df3%>%group_by(CONSULT_ORDER)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x = reorder(CONSULT_ORDER,-count),y = count),color='blue',fill='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 6: Frequency of factors in consult_order feature')
```
<br>

### View consult_charge feature
```{r,warning=FALSE}
df3%>%group_by(CONSULT_CHARGE)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x = reorder(CONSULT_CHARGE,-count),y = count),color='blue',fill='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 7: Frequency of factors in consult_charge feature')
```
<br>

### View Weekday_arr,weekday_dep features (14)
```{r,warning=FALSE}
a=df3%>%group_by(WEEKDAY_ARR)%>%summarise(count=n())%>%data.frame()
b=df3%>%group_by(WEEKDAY_DEP)%>%summarise(count2=n())%>%data.frame()
c=data.frame(cbind(a,b[,2]))
colnames(c)=c('Weekday','count_arr','count_dep')
c

data.m <- melt(c, id.vars='Weekday')
data.m%>%ggplot()+geom_col(aes(x = Weekday,y = value, fill=variable),position = "dodge")+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 8: Frequency of factors in weekday_arr and weekday_dep features')

a=table(df3$WEEKDAY_ARR,df3$WEEKDAY_DEP)

kable(a,caption = 'Table 5b: Relation between weekday_arr and weekday_dep')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
<br>


### View month_arr,month_dep features
```{r,warning=FALSE}
a=df3%>%group_by(MONTH_ARR)%>%summarise(count=n())%>%data.frame()
colnames(a)=c('Month','MONTH_ARR_Count')
b=df3%>%group_by(MONTH_DEP)%>%summarise(count2=n())%>%data.frame()
colnames(b)=c('Month','MONTH_DEP_Count')
c=full_join(a,b)

data.m <- melt(c, id.vars='Month')
data.m%>%ggplot()+geom_col(aes(x = Month,y = value, fill=variable),position = "dodge")+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 9: Frequency of factors in month_arr and month_dep features')+
  scale_x_continuous(name="Month", breaks = seq(1,12,1))

a=table(df3$MONTH_ARR,df3$MONTH_DEP)

kable(a,caption = 'Table 5c: Relation between month_arr and month_dep')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```
<br>


### View diag_details feature
```{r,warning=FALSE}
df3%>%group_by(DIAG_DETAILS)%>%summarise(count=n())%>%
  ggplot()+geom_col(aes(x = DIAG_DETAILS,y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 10: Frequency of observations for each value of diag_details feature')
```
<br>


### View charges feature
```{r,warning=FALSE}
df3$CHARGES=as.numeric(as.character(df3$CHARGES))
df3%>%ggplot()+geom_histogram(aes(x=CHARGES,y = ..density..),fill='red',color='red')+
  ylab('Density of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 11: Density of observations for each value of charge feature')

df3[is.na(df3$CHARGES),]%>%group_by(FINANCIAL_CLASS)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x=reorder(FINANCIAL_CLASS,-count),y=count),fill='blue',color='blue')+
  ylab('Number of observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 12: Number of observations with missing value of charge by financial_class')

df3[is.na(df3$CHARGES),]%>%group_by(ED_RESULT)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x=reorder(ED_RESULT,-count),y=count),fill='blue',color='blue')+
  ylab('Number of observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 13: Number of observations with missing value of charge by ed_result')

df3%>%group_by(FINANCIAL_CLASS)%>%summarise(mean_charge=mean(CHARGES,na.rm = TRUE))%>%ggplot()+
  geom_col(aes(x=reorder(FINANCIAL_CLASS,-mean_charge),y=mean_charge),fill='blue',color='blue')+
  ylab('Mean Value of Charge')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 14: Mean charge by financial_class')

df3%>%group_by(ED_RESULT)%>%summarise(mean_charge=mean(CHARGES,na.rm = TRUE))%>%ggplot()+
  geom_col(aes(x=reorder(ED_RESULT,-mean_charge),y=mean_charge),fill='blue',color='blue')+
  ylab('Mean Value of Charge')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 15: Mean charge by factors in ed_result feature')
```
<br>

### View ed_result feature
```{r,warning=FALSE}
df3%>%group_by(ED_RESULT)%>%summarise(count=n())%>%
  ggplot()+geom_col(aes(x = reorder(ED_RESULT,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 16: Frequency of observations for each value of ed_result feature')
```
<br>


### View hour_arr and hour_dep features
```{r,warning=FALSE}
a=df3%>%group_by(HOUR_ARR)%>%summarise(count=n())%>%data.frame()
colnames(a)=c('Hour','Arr_count')
b=df3%>%group_by(HOUR_DEP)%>%summarise(count=n())%>%data.frame()
colnames(b)=c('Hour','Dep_count')
c=full_join(a,b)

data.m <- melt(c, id.vars='Hour')
data.m%>%ggplot()+geom_col(aes(x = Hour,y = value, fill=variable),position = "dodge")+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 17: Number of observation for each factor in hour_arr and hour_dep features')+
  scale_x_continuous(name="Month", breaks = seq(1,24,1))

a=table(df3$HOUR_ARR,df3$HOUR_DEP)

kable(a,caption = 'Table 5d: Relation between hour_arr and hour_dep')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```
<br>


### View age feature
```{r,warning=FALSE}
df3%>%ggplot()+geom_histogram(aes(x=AGE,y = ..density..),fill='blue',color='blue')+
  ylab('Density of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 18: Density of observations for each value of age feature')

a=df3%>%mutate(age_category=cut(AGE,breaks = c(0,17,40,65,99,Inf),labels = c('<=17','18_40','40_65','66_99','>=100')))%>%
  group_by(age_category)%>%summarise(count=n())

kable(a,caption = 'Table 5: Count of observations by age category')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
<br>


### View gender feature
```{r,warning=FALSE}
df3%>%group_by(GENDER)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x=GENDER,y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 19: Number of observations for each gender')
```
<br>


### View hospital feature (22)
```{r,warning=FALSE}
df3%>%group_by(HOSPITAL)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x=reorder(HOSPITAL,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 20: Number of observations for each hospital')
```
<br>


### View race and ethnicity feature
```{r,warning=FALSE}
df3%>%group_by(RACE)%>%summarise(count=n())%>%ggplot()+
  geom_col(aes(x=reorder(RACE,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 21: Number of observations for each race')


df3%>%group_by(ETHNICITY)%>%summarise(count=n())%>%
  ggplot()+geom_col(aes(x=reorder(ETHNICITY,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 0))+
  ggtitle('Figure 22: Number of observations for each ethnicity')

a=df3%>%dplyr::select(RACE,ETHNICITY)%>%group_by(RACE,ETHNICITY)%>%summarise(count=n())

kable(a,caption = 'Table 6: Race and ethnicity matchup')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))

df3[(df3$RACE=='Asian' & df3$ETHNICITY==''),]$ETHNICITY='Not Hispanic or Latino'
df3[(df3$RACE=='Black or African American' & df3$ETHNICITY==''),]$ETHNICITY='Not Hispanic or Latino'
df3[(df3$RACE=='White' & df3$ETHNICITY==''),]$ETHNICITY='Not Hispanic or Latino'
df3[(df3$RACE=='Unknown' & df3$ETHNICITY==''),]$ETHNICITY='Unknown'
df3[(df3$RACE=='Native Hawaiian or Other Pacific Islander' & df3$ETHNICITY==''),]$ETHNICITY='Not Hispanic or Latino'
df3[(df3$RACE=='American Indian or Alaskan Native' & df3$ETHNICITY==''),]$ETHNICITY='Not Hispanic or Latino'
```
<br>

### View financial_class feature
```{r,warning=FALSE}
df3%>%group_by(FINANCIAL_CLASS)%>%summarise(count=n())%>%
  ggplot()+geom_col(aes(x=reorder(FINANCIAL_CLASS,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 23: Number of observations for each financial class')
```
<br>

### View dc_result feature
```{r,warning=FALSE,fig.width=6, fig.height=10}
df3%>%group_by(DC_RESULT)%>%summarise(count=n())%>%
  ggplot()+geom_col(aes(x=reorder(DC_RESULT,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 24: Number of observations for each factor in dc_result feature')
```
<br>

### View diagnosis feature
```{r,warning=FALSE}
df3%>%group_by(DIAGNOSIS)%>%summarise(count=n())%>%
  ggplot()+geom_col(aes(x=reorder(DIAGNOSIS,-count),y = count),fill='blue',color='blue')+
  ylab('Number of Observations')+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 25: Number of observations for each factor in diagnosis')
```
<br>


## Split data back into train and test set
```{r,warning=FALSE}
dtrain=df3[1:38221,]
dtrain=data.frame(cbind(dtrain,df[,27]))
colnames(dtrain)[27]='RETURN'
dtest=df3[38222:50253,]
```
<br>



### Fix issues related to missing entries, and data type
```{r,warning=FALSE}

dtrain$INDEX=NULL
dtrain$CONSULT_IN_ED[is.na(dtrain$CONSULT_IN_ED)]=0
dtrain$CONSULT_IN_ED=as.factor(dtrain$CONSULT_IN_ED)

dtrain$RISK=as.character(dtrain$RISK)
dtrain$RISK[dtrain$RISK=='']='Not Measured'
dtrain$RISK=as.factor(dtrain$RISK)


dtrain$SEVERITY=as.character(dtrain$SEVERITY)
dtrain$SEVERITY[dtrain$SEVERITY=='']='Not Measured'
dtrain$SEVERITY=as.factor(dtrain$SEVERITY)


dtrain$ADMIT_RESULT=as.character(dtrain$ADMIT_RESULT)
dtrain$ADMIT_RESULT[dtrain$ADMIT_RESULT=='']='Others'
dtrain$ADMIT_RESULT=as.factor(dtrain$ADMIT_RESULT)

dtrain$ACUITY_ARR=as.character(dtrain$ACUITY_ARR)
dtrain=dtrain[!(dtrain$ACUITY_ARR=='5 Purple'),]
dtrain$ACUITY_ARR=as.factor(dtrain$ACUITY_ARR)


dtrain$SAME_DAY=as.factor(dtrain$SAME_DAY)
dtrain$CONSULT_ORDER=as.factor(dtrain$CONSULT_ORDER)
dtrain$CONSULT_CHARGE=as.factor(dtrain$CONSULT_CHARGE)


#dtrain$WEEKDAY_ARR=ifelse((dtrain$WEEKDAY_ARR==1 | dtrain$WEEKDAY_ARR==7),'Weekend','Weekday')
dtrain$WEEKDAY_ARR=as.factor(dtrain$WEEKDAY_ARR)
dtrain$WEEKDAY_DEP=NULL


#for (i in 1:nrow(dtrain)) {
#  if(dtrain$MONTH_ARR[i]=='12'|dtrain$MONTH_ARR[i]=='1' |dtrain$MONTH_ARR[i]=='2'){
#    dtrain$MONTH_ARR[i]='Winter'
#  }
#  else if(dtrain$MONTH_ARR[i]=='3' |dtrain$MONTH_ARR[i]=='4' |dtrain$MONTH_ARR[i]=='5') {
#    dtrain$MONTH_ARR[i]='Spring'
#  }
#  else if(dtrain$MONTH_ARR[i]=='6' |dtrain$MONTH_ARR[i]=='7' |dtrain$MONTH_ARR[i]=='8'){
#    dtrain$MONTH_ARR[i]='Summer'
#  }
#  else{
#    dtrain$MONTH_ARR[i]='Autumn'
#  }
#}
dtrain$MONTH_ARR=as.factor(dtrain$MONTH_ARR)


#for (i in 1:nrow(dtrain)) {
#  if(dtrain$MONTH_DEP[i]=='12'|dtrain$MONTH_DEP[i]=='1' |dtrain$MONTH_DEP[i]=='2'){
#    dtrain$MONTH_DEP[i]='Winter'
#  }
#  else if(dtrain$MONTH_DEP[i]=='3' |dtrain$MONTH_DEP[i]=='4' |dtrain$MONTH_DEP[i]=='5') {
#    dtrain$MONTH_DEP[i]='Spring'
#  }
#  else if(dtrain$MONTH_DEP[i]=='6' |dtrain$MONTH_DEP[i]=='7' |dtrain$MONTH_DEP[i]=='8'){
#    dtrain$MONTH_DEP[i]='Summer'
#  }
#  else{
#    dtrain$MONTH_DEP[i]='Autumn'
#  }
#}
dtrain$MONTH_DEP=as.factor(dtrain$MONTH_DEP)



#dtrain$HOUR_ARR=ifelse((dtrain$HOUR_ARR<18 | dtrain$HOUR_ARR<8),'Daytime','Evening')
#dtrain$HOUR_DEP=ifelse((dtrain$HOUR_DEP<18 | dtrain$HOUR_DEP<8),'Daytime','Evening')
dtrain$HOUR_DEP=NULL

dtrain$CHARGES=as.numeric(as.character(dtrain$CHARGES))
#dtrain$CHARGES[is.na(dtrain$CHARGES) & dtrain$ED_RESULT=='Arrived in Error']=0
#dtrain$CHARGES[is.na(dtrain$CHARGES)]=median(dtrain$CHARGES,na.rm = TRUE)

dtrain$ED_RESULT=as.character(dtrain$ED_RESULT)
#dtrain$ED_RESULT[dtrain$ED_RESULT=='']='Discharge'
dtrain$ED_RESULT=as.factor(dtrain$ED_RESULT)

#dtrain=dtrain[!(dtrain$GENDER==''),]


#dtrain$RACE[dtrain$RACE=='']='Black or African American'
#dtrain$RACE[dtrain$RACE=='']='Black or African American'
#dtrain$ETHNICITY[dtrain$ETHNICITY=='']='Not Hispanic or Latino'

dtrain$DC_RESULT=as.character(dtrain$DC_RESULT)
#dtrain$DC_RESULT[dtrain$DC_RESULT=='']='Home or Self Care'
#dtrain$DC_RESULT=ifelse(dtrain$DC_RESULT=='Home or Self Care','Home or Self Care','Others')
dtrain$DC_RESULT=as.factor(dtrain$DC_RESULT)


dtrain[dtrain=='']=NA
dtrain[dtrain=='#N/A']=NA
dtrain[dtrain=='#VALUE!']=NA
dtrain=na.omit(dtrain)
dtrain=droplevels(dtrain)
```
<br>



## Exploratory analysis of training dataset
<br>

### View by hospitals and returns
```{r,warning=FALSE}
dtrain%>%group_by(HOSPITAL,RETURN)%>%summarise(count=n())%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x=HOSPITAL,y=Percent,fill=RETURN))+
  ggtitle('Figure 26: Return vs hospital')
```
<br>

### View by gender and returns
```{r,warning=FALSE}
dtrain%>%group_by(GENDER,RETURN)%>%summarise(count=n())%>%group_by(GENDER)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x=GENDER,y=Percent,fill=RETURN))+
  ggtitle('Figure 27: Return vs gender')
```
<br>

### View by age and returns
```{r,warning=FALSE}
dtrain%>%ggplot()+geom_boxplot(aes(x = RETURN,y = AGE))+ylab('Density of Observations')+
  ggtitle('Figure 28: Return vs age')
```
<br>

### View by race and return
```{r,warning=FALSE}
dtrain%>%group_by(RACE,RETURN)%>%summarise(count=n())%>%
  group_by(RACE)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x=RACE,y=Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 29: Return vs race')
```
<br>

### View by ethnicity and return
```{r,warning=FALSE}
dtrain%>%group_by(ETHNICITY,RETURN)%>%summarise(count=n())%>%
  group_by(ETHNICITY)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = ETHNICITY,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 30: Return vs ethnicity')
```
<br>

### View by financial class and return
```{r,warning=FALSE}
dtrain%>%group_by(FINANCIAL_CLASS,RETURN)%>%summarise(count=n())%>%
  group_by(FINANCIAL_CLASS)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = FINANCIAL_CLASS,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 31: Return vs hospital')
```
<br>

### View by weekday arrival and return
```{r,warning=FALSE}
dtrain%>%group_by(WEEKDAY_ARR,RETURN)%>%summarise(count=n())%>%group_by(WEEKDAY_ARR)%>%
  mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = WEEKDAY_ARR,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 32: Return vs weekday arrival')
```
<br>

### View by hour arrival and return
```{r,warning=FALSE}
dtrain%>%ggplot()+geom_boxplot(aes(x=RETURN,y = HOUR_ARR))+xlab('Returning Customers')+
  ggtitle('Figure 33: Return vs hour arrival')
```
<br>

### View by month arrival and return
```{r,warning=FALSE}
dtrain%>%group_by(MONTH_ARR,RETURN)%>%summarise(count=n())%>%group_by(MONTH_ARR)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = MONTH_ARR,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 34: Return vs month arrival')
```
<br>


### View by month departure and return
```{r,warning=FALSE}
dtrain%>%group_by(MONTH_DEP,RETURN)%>%summarise(count=n())%>%group_by(MONTH_DEP)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+
  geom_col(aes(x = MONTH_DEP,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 37: Return vs month departure')
```
<br>

### View by same day and return
```{r,warning=FALSE}
dtrain%>%group_by(SAME_DAY,RETURN)%>%summarise(count=n())%>%group_by(SAME_DAY)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+
  geom_col(aes(x = SAME_DAY,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 38: Return vs same day')
```
<br>

### View by ed_result and return
```{r,warning=FALSE}
dtrain%>%group_by(ED_RESULT,RETURN)%>%summarise(count=n())%>%group_by(ED_RESULT)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+
  geom_col(aes(x = ED_RESULT,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 39: Return vs ed_result')
```
<br>

### View by acuity of arrival and return
```{r,warning=FALSE}
dtrain%>%group_by(ACUITY_ARR,RETURN)%>%summarise(count=n())%>%group_by(ACUITY_ARR)%>%mutate(Percent=count/sum(count))%>%
  ggplot+geom_col(aes(x=ACUITY_ARR,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 40: Return vs acuity_arr')
```
<br>

### View by dc_result and return
```{r,warning=FALSE}
dtrain%>%group_by(DC_RESULT,RETURN)%>%summarise(count=n())%>%
  group_by(DC_RESULT)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = DC_RESULT,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 41: Return vs dc_result')
```
<br>

### View by admit_result and return
```{r,warning=FALSE}
dtrain%>%group_by(ADMIT_RESULT,RETURN)%>%summarise(count=n())%>%
  group_by(ADMIT_RESULT)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = ADMIT_RESULT,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 42: Return vs admit_result')
```
<br>

### View by consult_order and return
```{r,warning=FALSE}
dtrain%>%group_by(CONSULT_ORDER,RETURN)%>%summarise(count=n())%>%group_by(CONSULT_ORDER)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+
  geom_col(aes(x = CONSULT_ORDER,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 43: Return vs consult_order')
```
<br>

### View by consult_charge and return
```{r,warning=FALSE}
dtrain%>%group_by(CONSULT_CHARGE,RETURN)%>%summarise(count=n())%>%group_by(CONSULT_CHARGE)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+
  geom_col(aes(x = CONSULT_CHARGE,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 44: Return vs consult_charge')
```
<br>

### View by consult_in_ed and return
```{r,warning=FALSE}
dtrain%>%group_by(CONSULT_IN_ED,RETURN)%>%summarise(count=n())%>%group_by(CONSULT_IN_ED)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+
  geom_col(aes(x = CONSULT_IN_ED,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 45: Return vs consult_in_ed')
```
<br>

### View by diagnosis and return
```{r,warning=FALSE}
dtrain%>%group_by(DIAGNOSIS,RETURN)%>%summarise(count=n())%>%group_by(DIAGNOSIS)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = DIAGNOSIS,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 46: Return vs diagnosis')
```
<br>

### View by diag_details and return
```{r,warning=FALSE}
dtrain%>%group_by(DIAG_DETAILS,RETURN)%>%summarise(count=n())%>%group_by(DIAG_DETAILS)%>%
  mutate(Percent=count/sum(count))%>%ggplot()+geom_col(aes(x = DIAG_DETAILS,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 44: Return vs diag_details')
```
<br>

### View by risk and return
```{r,warning=FALSE}
dtrain%>%group_by(RISK,RETURN)%>%summarise(count=n())%>%group_by(RISK)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = RISK,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 47: Return vs risk')
```
<br>

### View by severity and return
```{r,warning=FALSE}
dtrain%>%group_by(SEVERITY,RETURN)%>%summarise(count=n())%>%group_by(SEVERITY)%>%mutate(Percent=count/sum(count))%>%
  ggplot()+geom_col(aes(x = SEVERITY,y = Percent,fill=RETURN))+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Figure 48: Return vs severity')
```
<br>

### View by CHARGES
```{r,warning=FALSE}
dtrain%>%ggplot()+geom_boxplot(aes(x=RETURN,y = CHARGES))+xlab('Returning Customers')+
  ggtitle('Figure 49: Return vs charges')
```


## Inferential Model
<br>

### Logsitic regression on full training dataset
```{r,warning=FALSE}
model.lm=glm(RETURN~.,data = dtrain,family = 'binomial')
summary(model.lm)
```
<br>

### Stepwise Logsitic regression on full training dataset
```{r,warning=FALSE}
model_all=glm(RETURN~.,data = dtrain,family = 'binomial')
model_null=glm(RETURN~1,data = dtrain,family = 'binomial')
a=step(object = model_null,scope = list(upper=model_all),direction = 'both',trace = 0)
summary(a)
```
<br>

## Predictive Model
<br>

### Split dtrain dataset into train and test set 
```{r,warning=FALSE}
set.seed(1234)

dtrain$DC_RESULT=ifelse(dtrain$DC_RESULT=='Home or Self Care','Home or Self Care','Others')
dtrain$DC_RESULT=as.factor(dtrain$DC_RESULT)

index=sample.split(dtrain$RETURN,SplitRatio = 0.75)
dtrain_train=dtrain[index,]
dtrain_test=dtrain[!index,]

set.seed(4567)
index2=sample.split(dtrain_train$RETURN,SplitRatio = 0.75)
dtrain_valid=dtrain_train[!index2,]
dtrain_train2=dtrain_train[index2,]
```
<br>


### Baseline Accuracy
```{r,warning=FALSE}
a=table(dtrain_test$RETURN)

kable(a,caption = 'Table 7: Matrix for Baseline Model')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Actual Values" = 1))

acc_b=max(a)/sum(a)
print(paste('The accuracy of the baseline model at 0.5 cutoff is:',acc_b))
```
<br>

### Logistic regression all variables
```{r,warning=FALSE}
model.lm=glm(RETURN~.,data = dtrain_train,family = 'binomial')
pred=predict(model.lm,dtrain_test,type = 'response')
pred=ifelse(pred>0.5,'Yes','No')
cm=table(dtrain_test$RETURN,pred,dnn = c('Actual','Predicted'))

kable(cm,caption = 'Table 9: Confusion Matrix for logistic regression -stepwise')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))

acc_lm=(cm[1,1]+cm[2,2])/sum(cm)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_lm))
```
<br>

### Logistic regression (Stepwise)
```{r,warning=FALSE}
model_all=glm(RETURN~.,data = dtrain_train,family = 'binomial')
model_null=glm(RETURN~1,data = dtrain_train,family = 'binomial')
a=step(object = model_null,scope = list(upper=model_all),direction = 'both',trace = 0)
pred=predict(a,dtrain_test,type = 'response')
pred=ifelse(pred>0.5,'Yes','No')
cm=table(dtrain_test$RETURN,pred,dnn = c('Actual','Predicted'))

kable(cm,caption = 'Table 9: Confusion Matrix for logistic regression -stepwise')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))

acc_lm=(cm[1,1]+cm[2,2])/sum(cm)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_lm))
```
<br>




### Classification Tree
```{r,warning=FALSE}
model_tree=tree(RETURN~.,data = dtrain_train)
pred=predict(model_tree,newdata = dtrain_test)
pred=ifelse(pred[,1]>0,1,0)

cm=table(pred,dtrain_test$RETURN)

kable(cm,caption = 'Table 12: Confusion Matrix for classification tree')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 1))

acc_ct=(cm[1,1])/sum(cm)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_ct))
```
<br>

### Split dtrain dataset into train and test set 
```{r,warning=FALSE}
set.seed(1234)

dtrain$DC_RESULT=ifelse(dtrain$DC_RESULT=='Home or Self Care','Home or Self Care','Others')
dtrain$DC_RESULT=as.factor(dtrain$DC_RESULT)

dtrain2=model.matrix(~.-1,data = dtrain)%>%data.frame()
index=sample.split(dtrain2$RETURNYes,SplitRatio = 0.75)
dtrain_train=dtrain2[index,]
dtrain_test=dtrain2[!index,]

set.seed(4567)
index2=sample.split(dtrain_train$RETURNYes,SplitRatio = 0.75)
dtrain_valid=dtrain_train[!index2,]
dtrain_train2=dtrain_train[index2,]
```
<br>


### LDA
```{r,warning=FALSE}
model_lda=lda(RETURNYes~.,data = dtrain_train)
pred=predict(model_lda,newdata = dtrain_test)

cm=table(pred$class,dtrain_test$RETURNYes)

kable(cm,caption = 'Table 13: Confusion Matrix for linear discriminant analysis')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))

acc_lda=(cm[1,1])/sum(cm)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_lda))
```
<br>


### XGboost - For phase II
```{r,warning=FALSE}
l=ncol(dtrain_train)
X_train<-as.matrix(dtrain_train[,1:(l-1)])
Y_train<-dtrain_train[,l]
train_matrix<-xgb.DMatrix(data=X_train, label=Y_train)
xgb_params <- list(booster="gbtree",
                   objective = "binary:logistic",
                   eval_metric = "auc",
                   eta=0.1,
                   subsample=1,
                   max_depth=100,
                   alpha=1,
                   lambda=1,
                   gamma=1,
                   min_child_weight=20,
                   max_delta_step = 20,
                   colsample_bytree=1)

bst<-xgboost(params = xgb_params,data=X_train,label =Y_train,nrounds = 33,verbose = FALSE)
X_test<-as.matrix(dtrain_test[,1:(l-1)])
X_test_matrix<-xgb.DMatrix(data=X_test)
dtrain_test_pred<-predict(bst,X_test_matrix)
test_pred=ifelse(dtrain_test_pred>0.5,'Yes','No')
cm_xgb=table(dtrain_test$RETURNYes,test_pred,dnn = c('Actual','Predicted'))
kable(cm_xgb,caption = 'Table 10: Confusion Matrix for Extreme Gradient Boosting Model')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))
acc_xgb=(cm_xgb[1,1]+cm_xgb[2,2])/sum(cm_xgb)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_xgb))
```
<br>

### Random Forest - For Phase II
```{r,warning=FALSE}
dtrain_train$RETURNYes=as.factor(dtrain_train$RETURNYes)
model.rf=ranger(RETURNYes~.,data = dtrain_train,
                mtry = 49,
                num.trees = 500,
                min.node.size = 30,
                sample.fraction = 0.65,
                seed = 123,
                verbose = FALSE)

a=predict(model.rf,dtrain_test)
cm_rf=table(dtrain_test$RETURNYes,a$predictions)
kable(cm_rf,caption = 'Table 11: Confusion Matrix for Random Forests')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))
acc_rf=(cm_rf[1,1]+cm_rf[2,2])/sum(cm_rf)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_rf))
```
<br>

### QDA
```{r,warning=FALSE}
set.seed(1234)

dtrain2=dtrain

dtrain2$ED_RESULT=as.character(dtrain2$ED_RESULT)
dtrain2$ED_RESULT=ifelse(dtrain2$ED_RESULT=='Discharge','Discharge',
                        (ifelse(dtrain2$ED_RESULT=='Admit','Admit','Others')))
dtrain2$ED_RESULT=as.factor(dtrain2$ED_RESULT)

dtrain2$RACE=as.character(dtrain2$RACE)
dtrain2$RACE=ifelse(dtrain2$RACE=='Black or African American','Black or African American',
                    ifelse(dtrain2$RACE=='White','White','Others'))
dtrain2$RACE=as.factor(dtrain2$RACE)
 

dtrain2$SEVERITY=NULL

index=sample.split(dtrain2$RETURN,SplitRatio = 0.75)
dtrain_train=dtrain2[index,]
dtrain_test=dtrain2[!index,]

set.seed(4567)
index2=sample.split(dtrain_train$RETURN,SplitRatio = 0.75)
dtrain_valid=dtrain_train[!index2,]
dtrain_train2=dtrain_train[index2,]



model_lda=qda(RETURN~.,data = dtrain_train)
pred=predict(model_lda,newdata = dtrain_test)

cm=table(pred$class,dtrain_test$RETURN)

kable(cm,caption = 'Table 14: Confusion matrix for quadratic discriminant analysis')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))

acc_qda=(cm[1,1]+cm[2,2])/sum(cm)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_qda))
```
<br>


### KNN
```{r,warning=FALSE}
set.seed(1234)

df_knn=dtrain[,c(3,8,20,23,24)]
df_knn$RETURN=as.character(df_knn$RETURN)
df_knn$RETURN=ifelse(df_knn$RETURN=='Yes',1,0)

index=sample.split(df_knn$RETURN,SplitRatio = 0.75)
dtrain_train=df_knn[index,]
dtrain_test=df_knn[!index,]

set.seed(4567)
index2=sample.split(dtrain_train$RETURN,SplitRatio = 0.75)
dtrain_valid=dtrain_train[!index2,]
dtrain_train2=dtrain_train[index2,]


min_max=function(df){
  new_df=data.frame()
  col=ncol(df)
  for (cl in 1:col) {
    min=min(df[,cl])
    max=max(df[,cl])
    for (l in 1:nrow(df)) {
      new_df[l,cl]=(df[l,cl]-min)/(max-min)
    }
  }
  return(new_df)
}



train_X=dtrain_train[,c(1,2,3,4)]
s_train_X=dtrain_train2[,c(1,2,3,4)]
valid_X=dtrain_valid[,c(1,2,3,4)]
test_X=dtrain_test[,c(1,2,3,4)]


train_X=min_max(train_X)
s_train_X=min_max(s_train_X)
valid_X=min_max(valid_X)
test_X=min_max(test_X)

train_Y=dtrain_train[,c(5)]
s_train_Y=dtrain_train2[,c(5)]
valid_Y=dtrain_valid[,c(5)]
test_Y=dtrain_test[,c(5)]

df_valid=data.frame()
for (k in seq(30,44,2)) {
  pred_knn=knn(s_train_X,valid_X,s_train_Y,k = k)
  a=table(pred_knn,valid_Y)
  acc=(a[1,1]+a[2,2])/sum(a)
  df_valid[((k-30)/2+1),1]=k
  df_valid[((k-30)/2+1),2]=acc
}

k=df_valid[,1][df_valid[,2]==max(df_valid[,2])]

pred=knn(train_X,test_X,train_Y,k=k)
cm=table(pred,test_Y)

kable(cm,caption = 'Table 15: Confusion matrix for KNN Model')%>%
  kable_styling(bootstrap_options = c("striped", "hover"))%>%
  add_header_above(c(" ", "Predicted Values" = 2))

acc_knn=(cm[1,1]+cm[2,2])/sum(cm)
print(paste('The accuracy of the model at 0.5 cutoff is:',acc_knn))
```
<br>
